---
title: 'GIS 3 Lab 1'
author: "Ryan (Yigong) Wang" 
fontsize: 11pt
output:
  pdf_document: 
    latex_engine: xelatex
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages,amsmath,amssymb,float,amsfonts,enumitem,tcolorbox,bm,algorithm,fontspec,sansmath,fontenc,xcolor}
  \setmainfont{Helvetica}
geometry: margin=0.75in
fig_crop: no

---

```{r setup, include=FALSE}
library(sf)
library(raster)
library(knitr)
library(tidyverse)
#library(gridExtra)·
library(spData)
library(dplyr)
library(spDataLarge)
library(rgdal)
options(width=70, digits=6, scipen=8)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
# Set R output size a bit smaller than default
opts_chunk$set(size='small') 
# Set the default to NOT display R code
opts_chunk$set(echo=FALSE) 
```
This lab is aimed to practice R installation and some introductory materials of the course.\newline

# 1. R Version
First, we check the R version installed.
```{r, echo=FALSE}
version$version.string
```
As we can see, our R version is \underline{3.6.3}.

Next up, we will see some basic GIS examples achievable in R.

# 2. Simple Feature Geometries
The \textbf{sf} package contains multiple geometric construction functions that can be used to construct geometries such as points, lines, polygons and many more. This demonstration will showcase soome features of this functionality.

## Polygon With A Hole
One of the special geometries that \textbf{sf} is able to build is a polygon with a hole. The code chunk below will construct such a shape. It is done by creating a polygon and its hole and combining those two together. (Code is shown for this part.)
```{r, echo=TRUE}
polygon_border = rbind(c(1, 5), c(2, 2), c(4, 1), c(4, 4), c(1, 5))
polygon_hole = rbind(c(2, 4), c(3, 4), c(3, 3), c(2, 3), c(2, 4))
polygon_with_hole_list = list(polygon_border, polygon_hole)
plot(st_polygon(polygon_with_hole_list), col = sf.colors(12, categorical = TRUE), 
     border = 'grey', axes = TRUE)
```

```{r}
multipoint_matrix = rbind(c(5, 2), c(1, 3), c(3, 4), c(3, 2))

linestring_matrix = rbind(c(1, 5), c(4, 4), c(4, 1), c(2, 2), c(3, 2))

```

## Another Example
In this example, we draw two shapes together with the \textbf{st_geometrycollection} function. In this case, we combine a multipoint and a linestring:
```{r, fig.width=4, fig.height=4}
gemetrycollection_list = list(st_multipoint(multipoint_matrix),
                              st_linestring(linestring_matrix))
plot(st_geometrycollection(gemetrycollection_list), col = sf.colors(2, categorical = FALSE), border = 'grey', axes = FALSE, xlab=" ")
```

# 3. Basic Map Making Example
Using the \textbf{raster} package, we can make a map using the dataset in section 2.3 of the textbook. This map uses raster.
```{r, fig.width=5, fig.height=5, echo=TRUE}
raster_filepath = system.file("raster/srtm.tif", package = "spDataLarge")
new_raster = raster(raster_filepath)
plot(new_raster, axes=FALSE, )
```

This is the end of Lab 1 of GIS 3.

